---
title: "Table one"
author: "Duzhe Wang"
date: "11/25/2021"
output: html_document
---
## Introduction

- Most part of this note is from Chapter 1 of [Data Science for Biological, Medical and Health Research](https://thomaselove.github.io/432-notes/index.html). 

- Other reference: [Introduction to tableone](https://cran.r-project.org/web/packages/tableone/vignettes/introduction.html)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Building Table 1 for `fakestroke`: Attempt 1

```{r}
library(tidyverse)
library(tableone)
fakestroke=read_csv("fakestroke.csv")
```

```{r}
# generate a list of all of the variables in the tibble
dput(names(fakestroke))
```

```{r}
fs.vars=c("age", "sex", "nihss", "location", "hx.isch", 
"afib", "dm", "mrankin", "sbp", "iv.altep", "time.iv", "aspects", 
"ia.occlus", "extra.ica", "time.rand", "time.punc")

fs.trt=c("trt")

att1=CreateTableOne(data=fakestroke, 
                    vars=fs.vars,
                    strata=fs.trt)
print(att1)
```


## Buliding Table 1 for `fakestroke`: Attempt 2

Some of this is very helpful, and other parts need to be fixed:

- The 1/0 variables (`afib`, `dm`, `extra.ica`) might be better if they were treated as the factors they are, and reported as the Yes/No variables, with counts and percentages rather than means and sd. 

- In some cases, we may prefer to re-order the levels of the categorical variables, particularly the `mrankin` variable, but also the `ia.occlus` variable. It would also be more typicla to put the intervention group to th left and the control group to the right, so we may need to adjust our `trt` variable's levels accordingly. 

- For each of the quantitative variables, we should make a decision whether a summary with mean and sd is appropriate, or whether we should instead summarize with the median and quartiles. 

- If we'd left in the `time.punc` variable, we'd get some warnings, having to do with the fact that `time.punc` is only relevant to patients in the intervention group

```{r}
# specify the categorical variables
fs.factorvars=c("sex", "location", "hx.isch", "afib", "dm", "mrankin", "iv.altep", "ia.occlus", "extra.ica")

# re-order some of the categorical variabeles so that the levels are most useful to us
fakestroke=fakestroke%>%
  mutate(trt=fct_relevel(trt, "Intervention", "Control"),
         mrankin=fct_relevel(mrankin, "0", "1", "2", ">2"),
         ia.occlus=fct_relevel(ia.occlus, "Intracranial ICA", "ICA with M1", "M1", "M2", "A1 or A2")
         )
```




```{r}
att2=CreateTableOne(data=fakestroke,
                    vars=fs.vars,
                    factorVars = fs.factorvars,
                    strata=fs.trt)
print(att2)


```


```{r}
# specify certain quantitative variables as non-normal 
# specify exact tests for certain categorical variables 
print(att2, 
      nonnormal=c("age", "nihss",  "time.iv", "aspects", "time.rand"),
      exact=c("location", "mrankin"))


```



## Obtaining a more detailed summary

```{r}
summary(att2)


```


## Exporting the completed table 1 from R to Excel or Word

- Save and open in Excel

```{r}
fs.table1save=print(att2, 
                    nonnormal=c("age", "nihss", "time.iv", "aspects", "time.rand"),
                    exact=c("location", "mrankin"),
                    quote=FALSE, noSpaces=TRUE, printToggle=FALSE)
write.csv(fs.table1save, file="/Users/duzhewang/Desktop/tableone/fstable1.csv")


```




