---
title: "Normalization"
author: "Duzhe Wang"
date: "5/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Normalization is very important to improve the performance of machine learning algorithms. If we don't normalize the data, the machine learning algorithm will be dominated by the variables that use a larger scale, adversely affecting model performance. 

# Standardization 

In this tutorial, we use the `iris` dataset. We first introduce the standardization technique. Mathematically, standardization uses the following equation:

$$z=\frac{x-\mu}{\sigma},$$
where $\mu=\frac{1}{N}\sum_{i=1}^{N}x_i$ and $\sigma=\sqrt{\frac{1}{N-1}\sum_{i=1}^{N}(x_i-\mu)^2}$. 

- Bottom-up approach 

```{r}

standardization_vec=function(x){
  # x is a vector
  # this will make the new vector z has sample mean 0 and sample standard deviation 1
  
  mu=mean(x)
  sigma=sd(x)
  z=(x-mu)/sigma
  return(z)
}


standardization_df=function(dat){
  # dat is a data frame
  # the output is still a data frame 
  newdat=as.data.frame(lapply(dat, standardization_vec))
  return(newdat)
}


```

Now let's apply the above function to the iris data set. 

```{r}
iris_standard=standardization_df(iris[1:4])
summary(iris_standard)
mean(iris_standard$Sepal.Length)
sd(iris_standard$Sepal.Length)
```


- Standardization using `caret` pacakge 

```{r}
library(caret)

preproc=preProcess(iris[1:4], method=c("center", "scale"))
iris_standard_caret=predict(preproc, iris[1:4])
summary(iris_standard_caret)


```

- Standardization using `scale` function

```{r}
iris_standard_scale=as.data.frame(scale(iris[1:4]))
summary(iris_standard_scale)
```


# Min-max scaling 

Mathematically, min-max scaling is done as follows:

$$x_{norm}=\frac{x-x_{\min}}{x_{\max}-x_{\min}}$$


- Bottom-up approach 

```{r}
minmax_vec=function(x){
  x_minmax=(x-min(x))/(max(x)-min(x))
  return(x_minmax)
}


minmax_df=function(dat){
  newdat=as.data.frame(lapply(dat, minmax_vec))
  return(newdat)
}

```

Now let's see the iris example. 


```{r}
iris_minmax=minmax_df(iris[1:4])
summary(iris_minmax)
```


- Min-max using `caret` package

```{r}

preproc_minmax=preProcess(iris[1:4], method=c("range"))

iris_minmax_caret=predict(preproc_minmax, iris[1:4])

summary(iris_minmax_caret)
str(iris_minmax_caret)


```

# Reference

- [Normalizing data with R](https://www.pluralsight.com/guides/normalizing-data-r)







