---
title: "ROC curve"
author: "Duzhe Wang"
date: "5/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Goal 

The goal of this exercise is to plot the ROC curve and compute AUC. 

# 2. Basic concepts 

- The x-axis of ROC is false positive rate, which is defined as 

$$\text{FPR}=\frac{\text{false positive}}{\text{negative}}$$

- The y-axis of ROC is true positive rate, which is defined as 

$$\text{TPR}=\frac{\text{true positive}}{\text{positive}}$$

- AUC: area under the ROC curve

# 3. Code 

- Two classes: class 0 (positive) and class 1 (negative)

- ROC curve function: 

```{r}
roc=function(groundtruth, predprob, theta){
    # groudtruth: class 0 (positive) and class 1 (negative)
    # predprob: predicted probability 
    # theta: threshold between 0 and 1
    # false positive rate: FP/N  (FP: true negative, but predicted positive)
    # true positive rate (recall): TP/P
      
    groundtruth.negative=which(groundtruth==1)
    groundtruth.positive=which(groundtruth==0)
    
    # first column: theta; second column: FPR; third column: TPR
    rate=matrix(0, nrow=length(theta), ncol=3)   

    for (i in 1:length(theta)){
        # predicted class is 1 if predprob>theta; otherwise the predicted class is 0.
        predclass=as.integer(predprob>theta[i])  
        predclass.negative=which(predclass==1)
        predclass.positive=which(predclass==0)
        rate[i, 1]=theta[i]
        # false positive rate
        rate[i, 2]=length(intersect(groundtruth.negative,predclass.positive))/sum(groundtruth==1)      
        # true positive rate
        rate[i, 3]=length(intersect(groundtruth.positive,predclass.positive))/sum(groundtruth==0)      
     }
    
    # plot ROC 
    plot(rate[,2], rate[,3], xlab="FPR", ylab="TPR", type="l")
    
    # calculate AUC
    sortedfpr=sort(rate[,2], decreasing=FALSE)
    sortedtpr=sort(rate[,3], decreasing=FALSE)
    auc=sum((sortedfpr[2:length(theta)]-sortedfpr[1:(length(theta)-1)])*sortedtpr[1:(length(theta)-1)])
    
    # return output 
    return(list(rate=rate, auc=auc))
}
```


# 4. Examples 

```{r}
groundtruth=sample(c(0, 1), size=10000, replace=TRUE,prob = c(0.8,0.2))
predprob<-runif(10000,0,1)
theta=seq(from=0.05, to=1, by=0.001)
roc(groundtruth, predprob, theta)

```

**Explanation**: Let $n$ be the total sample size. Let $P(c=0)=p$ and $P(c=1)=1-p$. Since predprob is uniform, so given a fixed $\theta$, for $c=0$ group, there are around $np\theta$ observations which are predicted as 0 and around $np(1-\theta)$ observations which are predicted as 1. For $c=1$ group, there are around $n(1-p)\theta$ observations which are predicted as 0 and around $n(1-p)(1-\theta)$ observations which are predicted as 1. Therefore, the FPR is 

$$  \frac{n(1-p)\theta}{n(1-p)}=\theta,$$
and the TPR is 

$$\frac{np\theta}{np}=\theta.$$
Hence, the ROC curve is like the line $y=x$ and the AUC is 0.5. 














